---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  word_document: default
  html_document: default
  pdf_document: default
---
<center> <h1> Assignment 3 (10%) </h1> </center>
<center>  <h3> Ajay Herod </h2> </center>
<center> <h3> D10 500764104 </h2> </center>
---


#### 1. Import to R the following fiel: http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv (The dataset is related to white Portuguese "Vinho Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality) (3 points)
```{r}
#install.packages('RCurl')
library(RCurl)
web <- getURL("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv")
wine <- read.csv2(text = web)
```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
str(wine)
#OR
sapply(wine, class)
```
  
#### 3. Are there any missing values in the dataset? (4 points)
```{r}
sum(is.na(wine))
```

#### 4. What is the correlation between the attributes other than Quality? (10 points)
```{r}
wine <- as.data.frame(apply(wine, 2, as.numeric))
cor(wine[1:11])
```

#### 5. Graph the frequency distribution of wine quality by using Quality. (10 points)
```{r}
hist(wine$quality, xlab="Quality", main="Frequency Distribution of Wine Quality")
```

#### 6.	Reduce the levels of rating for quality to three levels as high, medium and low. Assign the levels of 3 and 4 to level 0; 5 and 6 to level 1; and 7,8 and 9 to level 2. (10 points)
```{r}
for (i in 1:length(wine$quality)){
  if(wine$quality[i] == 3 || wine$quality[i] == 4){
    wine$quality[i] = 0
  }
  else if(wine$quality[i] == 5 || wine$quality[i] == 6){
    wine$quality[i] = 1
  }
  else {
    wine$quality[i] = 2
  }
}
```

#### 7. Normalize the data set by using the following function: (12 points)
```{r}
normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
wine <- as.data.frame(lapply(wine, normalize))
summary(wine)
```

#### 8. Divide the dataset to training and test sets. (12 points)
```{r}
set.seed(1)
index <- sample(1:nrow(wine), 0.7*nrow(wine))
wine_train <- wine[index,]
wine_test <- wine[-index,]
```

#### 9. Use the KNN algorithm to predict the quality of wine using its attributes. (12 points)
```{r}
#install.packages("class")
library(class)
wine_train_labels <- wine_train[,12]
wine_test_labels <- wine_test[,12]
wine_test_pred <- knn(train = wine_train[,1:11],
                      test = wine_test[,1:11],
                      cl = wine_train_labels, k=10)
```

#### 10. Display the confusion matrix to evaluate the model performance. (12 points)
```{r}
table(Actual=wine_test_labels, Predicted=wine_test_pred)
```

#### 11. Evaluate the model performance by computing Accuracy, Sensitivity and Specificity. (12 points)
```{r}
#Accuracy = (TN+TP)/Total
(1019+152+1)/(1+41+1+1+1019+84+171+152)
#Sensitivity = TP/(TP+FN) For Actual 0
1/(1)
#Specificity = TN/(TN+FP) For Actual 0
(1019+84+171+152)/(1019+84+171+152+41+1)
#Sensitivity = TP/(TP+FN) For Actual 0.5
1019/(1019+41+171)
#Specificity = TN/(TN+FP) For Actual 0.5
(1+1+0+152)/(1+1+0+152+1+84)
#Sensitivity = TP/(TP+FN) For Actual 1
152/(152+84+1)
#Specificity = TN/(TN+FP) For Actual 1
(1+41+1+1019)/(1+41+1+1019+0+171)
```

This is the end of Assignment 3

Ceni Babaoglu, PhD